---
title: "Crime Rates"
---

Follow along with the demo to learn how to calculate crime rates using R and the Tidyverse. We'll also introduce a new library, janitor, which helps make column names easier to work with. You should get started by loading the tidyverse in a new R block, then install and load janitor.

### Task 1: Load the tidyverse and install and load janitor

```{r}
library(tidyverse)
#install.packages("janitor")
library(janitor)
```



### Task 2: Load the crime dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv

```{r}
crimes <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv")
```


### Task 3: Use janitor to clean up the column names

```{r}
crimes <- crimes |> clean_names()
```


### Task 4: Calculate the total crime rate for each year (picking our unit!)

```{r}
crimes |> mutate(total_rate = grand_total/population * 10000)
```



### Task 5: Calculate the violent and property crime rates for each year (same unit)

```{r}
crimes_with_rates <- crimes |> mutate(violent_rate = violent_crime_total/population * 10000, property_rate = property_crime_total/population * 10000)
```



### Task 6: Describe our results

Write a few sentences about what you found in the data. What years had the highest crime rates? What years had the lowest crime rates? Does the pattern match your expectations? What might explain some of the results you found?

The years with the highest violent crime rate were 1992, 1993 and 1995. At the time, crime was a very important part to the ongoing elections. The 90s were also a point in time where drug dealing and the crack cocaine situation were still very prevalent in neighborhoods. The years with the lowest violent and property crime rates were the late 2010s and early 2020s, largely because COVID-19 kept a lot of people in their houses. It is difficult to commit crimes if you have no incentive to leave the house. The year 1980 had the highest property crime rate, likely because of low amounts of supervision on teenage kids.



When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.

