---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

The data set is largely dominated by different types of assault, breaking and entering and robbery. However, the most incidents, in terms of the total number of them, lie within accidents, auto incidents and theft. From a reporting standpoint, I would say stolen vehicles immediately jumps out at me because it has some of the most total incidents and it also appears to have some spikes over the course of the year. Additionally, I would also say assault is a category that stands out because it constantly hovers in the 70s throughout the year. 

### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_totals <- pg_crime |> mutate(total_incidents=(`Jan`+`Feb`+`Mar`+`Apr`+`May`+`Jun`+`Jul`+`Aug`+`Sep`+`Oct`+`Nov`+`Dec`))
pg_crime_averages <- pg_crime_totals |> mutate(`per_month` = (total_incidents/12))
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

The overwhelming leader in average incidents per month is accidents with 481.3. Shortly below accidents are stolen automobiles (399.2) and theft from automobiles (391.9). The lowest average belongs to homicide at 5.8 incidents per month. The pattern of crime in Prince George's County largely revolves around vehicles, whether it's an accident on the road or some kind of stolen property from within a car. The average assaults per month was 70, which I thought was interesting because it stayed pretty consistently around that number each month of the year. So, one question I'd have is what is Prince George's County doing to help reduce that number since it seems relatively unchanged?


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(june_change=(`Jun`-`per_month`)/`per_month` * 100,
july_change=(`Jul`-`per_month`)/`per_month` * 100,
august_change=(`Aug`-`per_month`)/`per_month` * 100)
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

While there are some increases and decreases throughout the summer, I don't think there's enough evidence to claim that there is a trend associated with summer. Only three categories experienced a percentage increase in all three months: Homicide, theft from auto, and auto stolen. But other than that, I don't think there is sufficient enough evidence to conclude that crime is more likely to happen in the summer.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Car accidents were the most frequently reported incidents in Prince George's County, averaging 481.3 incidents per month over the course of 2025, according to the county's crime data.

Vehicle-related incidents made up four of the six categories of crime with the highest monthly average in cases. Automotive theft averaged 399.2 cases per month, thefts from automobiles averaged 391.9 cases per month and accidents with impounds averaged 320.8 cases per month. The next closest crime category was theft, which averaged 284.1 cases per month in 2025.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
