---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |> mutate(violent_rate_2019 = (violent2019 / pop2019) * 10000, property_rate_2019 = (property2019 / pop2019) * 10000, violent_rate_2020 = (violent2020 / pop2020) * 10000, property_rate_2020 = (property2020 / pop2020) * 10000)
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

The two highest cities in violent crime rate for both 2019 and 2020 were Baltimore and Ocean City, though they traded places at #1. A lot of that has to do with the fact that those are two of the most populated cities and population dense cities in the state. Ocean City also lead the property crime rates, which could be chalked up to it being a vacation town, which would leave a lot of belongings unattended. Cambridge's violent crime rate per 10,000 people spiked by almost 60.


### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |> mutate (violent_pct_change = ((violent_rate_2020 - violent_rate_2019) / violent_rate_2019) * 100, property_pct_change = ((property_rate_2020 - property_rate_2019) / property_rate_2019) * 100)
```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

I would say the patterns are more interesting and give more information for a story. The outliers in the percentage change are largely centered around cities and towns that have single-digit crimes in both categories. I think a potential story in the data is that Colmar Manor had nearly a 100% increase in property crime if a local angle were to be taken. Another potential story is that the state's two most popular cities, Baltimore and Frederick, both saw percentage decreases in property and violent crime.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Ocean City ranked among Maryland's top two cities in both violent and property crime rates for the second straight year, according to analysis of recent crime data in Maryland from 2019 and 2020.

The statistics show that the city posted a violent crime rate of 239.7 and property crime rate of 1212.8 per 10,000 people, both of which ranked first among Maryland cities for the year. The pattern is marked by a 82.9% increase in violent crime rate and a 17.0% property crime rate.

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.


### Task 10: Take the HW4_MD_CityCrime.Rmd notebook you worked on and add a section at the bottom called Task 10. Your task will be to take the violent crime rates for 2020 and do three things: Using the 2020 violent crime rate, calculate the percentiles using `ntile()`. Save the results to a new dataframe called `md_cities_percentiles`. How many cities are above the 90th percentile? 

```{r}
md_cities_percentiles <- md_cities_change |> mutate(percentile = ntile(violent_rate_2020, 100))

md_cities_percentiles |> summarize (
    violent_iqr_2020 = IQR(violent_rate_2020)
)
```

There are 11 cities in the 90th percentile or above.

The interquartile range for Maryland's violent crime rate in 2020 of 41.98895 does not necessarily change my story angle about Ocean City's overwhelmingly high crime rates. I would probably keep my story idea because the IQR is roughly 42 crimes per 100,000 residents and I don't think it is substantial enough to alter my story idea, which is about the high crime rates in Ocean City.